Complex mecanisms linking biomass stability and total biomass with species richness and network structure in stream ecosystems

### 2020-05-11 manuscript/bef_stability/manuscript.Rmd
Introduction done

#### Biodiversity and ecosystem functioning

Biodiversity is essential to determine and maintain ecosytem properties.
Current biodiversity losses question the ability of how ecosystems will be
affected in turn.

The relationship between biodiversity and ecosystem functioning has been a
central research topic in ecology [@may_will_1972; @hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @duffy_why_2009;
@isbell_forest_quantifying_2018].

@@

#### Shape of BEF: total biomass

Species richness has been reported to increase productivity
[@tilman_productivity_1996; @grace_integrative_2016] through complementarity
between species in resource exploitation and selection effects
[@loreau_partitioning_2001; @cardinale_species_2002; @fargione_selection_2007].
Empirical assessment of the effect of species richness on total biomass showed
negative [@grace_integrative_2016] and positive relationships
[@hector_plant_1999; @tilman_biodiversity_2014]. @grace_integrative_2016 showed
that the results depend on the refinement of the mecanisms involved in the
analysis.

#### Shape of BEF: stability

The relationship between species richness and stability is even more complex
partly because of the inherent multidimensionality of stability
[@donohue_dimensionality_2013; @donohue_navigating_2016] and because of the
different metrics of stability privileged in theoretical (stability of local
equilibrium, robustness to primary extinction) and empirical work (variability
of abundance or biomass through time). Empirical studies showed consistently
that species richness decreased community variability in grasslands
[@tilman_biodiversity_2006], mesocosms [@pennekamp_biodiversity_2018] and
vertebrates [@olivier_independent_nodate; @franssen_annual_2011]. Two mecanisms
are involved into community variability, (1) species variability and (2) the
species synchrony [@loreau_species_2008; @thibaut_understanding_2013].  Positive
[@tilman_biodiversity_2006], neutral [@jiang_different_2009;
@olivier_independent_nodate] and negative [@franssen_annual_2011;
@olivier_independent_nodate] relationships between species variability and
species richness have been found. On the other side, studies consistently found
that species richness decreased synchrony between species [@loreau_species_2008;
@thibaut_understanding_2013; @bluthgen_land_2016; @olivier_independent_nodate].
Beyond species richness, species interactions, which modifies species abondance
and biomass and make species dynamics interdependant, have been largely ignored
in empirical assessment.

#### BEF: Theoretical ecology

Conversely, theoretical ecology display an extensive literature about the
effects of network structure on stability [@may_will_1972;
@angelis_stability_1975; @stouffer_compartmentalization_2011;
@thebault_stability_2010; @barbier_pyramids_2019;
@shanafelt_david_w._stability_nodate].  Theoretical ecology generally uses
stability definitions related to community equilibrium and its stability to
small perturbations.  Connectance (i.e. proportion of realized links in a
network) can decrease [@angelis_stability_1975; @thebault_stability_2010] or
increase resilience of trophic network depending on the functioning of networks
[@angelis_stability_1975].  The number of trophic levels, the vertical position
in the trophic network as well as the strength of their biomass regulation (i.e.
top-down and bottom-up) can determine biomass and stability at equilibrium
[@hurd_stability_1971; @shanafelt_david_w._stability_nodate;
@zhao_horizontal_2019; @barbier_pyramids_2019].

A large inconsistency remains between theoretical and empirical assessments of
the relationship between community structure and ecosystem properties.
Theoretical studies have ignored temporal variability and then the relative
contribution of synchrony and species variability on community variability while
empirical studies about temporal variability and total biomass in natural
communities have ignored the contribution of network structure.

Our goal was to assess the mecanisms determining temporal variability and total
biomass in natural communities while integrating both species richness and
network structure. We combined together the effect of vertical and horizontal
diversity and thus aimed to bridge the current gap between theoretical and
empirical studies. We used an standardized monitoring of fish communities in
french streams. More than three hundreds sites were followed more than 10 years
on the 1995-2018 period. We inferred four thousands potential interaction
networks corresponding to each sampling event, based on trait matching.

We investigated the effects of the community structure on total biomass and
community variability, which in turn determine community variability. We have
also included the effect of environmental drivers [@grace_integrative_2016]
which can directly affect community variability or indirectly via community
structure.

- Q1: diversity-stability, diversity-biomass in aquatic systems
- Q2: Effect of horizontal and vertical diversity on total biomass and stablity
- Q3:

@@

We seeked to disantangle the unravel the links between
environment and community structure on both biomass stability and total biomass.
To investigate complex relationships, direct and indirect, between variables,
Structural Equation Modeling (SEM) is one promising way  
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. SEM makes possible to set causal relationships between
variables. This step can be critical, as the estimated parameters can be
determined by both the choice of the variables and the links between them.
Therefore, the set of variables and links shoud be supported by theory and
litterature.

We built a metamodel (set of variables and links between them). Basically, we
tested the links between environmental conditions, community structure and
ecosystem properties namely biomass stability on total biomass. Theory in
ecology predicts that community structure influences stability. In turn,
environment influences community structure but also directly stability.

>>>>

#### Biodiversity and ecosystem properties

Biodiversity is essential to determine and maintain ecosytem properties.
Current biodiversity losses question the ability of how ecosystems will be
affected in turn. Since 1970s, ecology aims to uncover the complex relationships
between biodiversity and ecosystem properties [@may_will_1972; @hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @duffy_why_2009;
@isbell_forest_quantifying_2018]. Because biodiversity largely determines in
turn both quantity and distribution of biomass or energy across ecosystems, it
has a tremendous effects on the properties exhibited by an ecosystem, such as
its size (total biomass) [REF], its production (production of biomass by units of
time) [REF] and the temporal stability of its biomass [REF].

#### Horizontal and vertical diversity are two ways to look at biodiversity

Biodiversity can be quantified in numerous ways, as the number of species, i.e.
species richness [REF] and in number of functional groups, i.e species are
gathered on what they do [REF]. Those metrics can be gathered into "horizontal
diversity", describing the list of components present in a community. An
orthogonal view, based on trophic chains, described species on what they eat. It 
can be defined as interactions or biomass/energy fluxes between species and how
they are distributed among species. The latter can be quantified as the number
of interaction between species or the number of trophic level. This orthogonal
view of seeing biodiversity, can be gathered into "vertical diversity".  
Horizontal and vertical diversity are two views of quantifying biodiversity that
are, however, not independent from each other [@zhao_horizontal_2019]. The add
or removal of one species can have different effects depending on its vertical
position in the trophic chain [REF duffy 2007]. Hence, it is crucial to
study their interplay to increase our understanding of ecosystem dynamics.

#### horizontal and vertical diversity increase total biomass by common and different mechanisms

Horizontal and vertical diversity has been shown to promote higher biomass
production in a variety of ecosystems. Horizontal diversity, as species richness, has been
reported to increase productivity through complementarity between species in
resource exploitation and selection effects in plant communities
[@tilman_productivity_1996; @cardinale_resource_2001; @grace_integrative_2016].
Regarding total biomass, empirical assessment have
shown generally positive relationships with species richness in plant
communities [@hector_plant_1999; @tilman_biodiversity_2014, but see
@grace_integrative_2016], as in mesocosms [@pennekamp_biodiversity_2018].
Theoretical models have also confirmed that an increase in horizontal diversity should
translate into increase in plant biomass  through
resource complementarity [@loreau_biodiversity_1998] and selection effects [@tilman_plant_1997]. The
addition of trophic levels, notably consumer on plants generally breaks the
positive relationship between plant biomass and plant diversity as a significant
amount of biomass is transfered to consumers. It has been shown both in
experiments [REF] and theoretical models [@thebault_food-web_2003]. Futhermore,
at higher trophic level, the resource complementary of consumers can increase as
the number of consumer species increase [@thebault_food-web_2003]. However, the
joint effects of horizontal and vertical diversity on total biomass has been
rarely tested. It is not clear whether horizontal and vertical diversity just
change the distribution of biomass across trophic level or it increases the
total amount of  biomass.

One reason might be the focus on plant communities, which are relatively easy to
manipulate and have an important role on determining the quantity and quality of
biomass for the other compartments of ecosystems [REF]. One trophic level is
often considered, and so the comp

@@

#### Horizontal and vertical diversity stability

Horizontal and vertical diversity also affects the stability of ecosytem
properties. Experimental studies showed consistently that species richness
increase temporal biomass stability in grasslands [@tilman_biodiversity_2006],
mesocosms [@pennekamp_biodiversity_2018]. While in natural communities,
same relation has been found for temporal biomass stability in stream fish communities 
[@franssen_annual_2011] and temporal abundance stability of birds and butterfies
[@olivier_independent_nodate, but not bats]. Two mecanisms are involved into
temporal stability of communities, (1) synchrony and (2) variability
[@loreau_species_2008; @thibaut_understanding_2013]. Synchrony mesures the
covariation of species through time. Very low synchrony indicates compensatory
temporal variation between species and so increased temporal stability. studies
consistently found that species richness decreased synchrony between species
[@loreau_species_2008; @thibaut_understanding_2013; @bluthgen_land_2016;
@olivier_independent_nodate]. (*Should I add more details about the
mecanisms?*). The average temporal variability of species is measured as a
average coefficient of variation of species, weighted by species
abundance/biomass. Higher species temporal variabiliy thus results in lower
temporal stability. Positive [@tilman_biodiversity_2006], neutral
[@jiang_different_2009; @olivier_independent_nodate] and negative
[@franssen_annual_2011; @olivier_independent_nodate] relationships between
species variability and species richness have been found. Hence, synchrony
has been found to be the main driver of temporal
stability [@tilman_biodiversity_2006; @olivier_independent_nodate] and
responsible for the positive relationship between horizontal diversity and
temporal stability of biomass and abundance.

Vertical diversity, represented by network structure, has been central to the
the literature of diversity-stability debate and largely explored by theoretical
ecology [@may_will_1972; @angelis_stability_1975;
@stouffer_compartmentalization_2011; @thebault_stability_2010;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate]. Because species
interactions make species dynamics interdependant, they have a tremendous role
on stability notably through its effects on the propagation of perturbation.
However, it is important to note that theoretical ecology generally uses
stability definitions related to community equilibrium and its stability to
small perturbations [@may_will_1972; @angelis_stability_1975;
@barbier_pyramids_2019; @shanafelt_david_w._stability_nodate] or to the number
of secondary extinctions of perturbation [@duffy_why_2009]. Those stability 
metrics harly relates to empirical temporal stability [see
@arnoldi_resilience_2016]. However, the work of @thebault_trophic_2005 shows
that asynchrony consistenly increase temporal stability and the positive
relationship between temporal stability and species richness. However asynchrony
can also increase species variability through interspecic competition, which
amplifies species fluctuations. Furthermore species interactions, considered as 
as competition strength and niche breath of consumers can results in negative or
positive temporal stability-species richness relationships
[@thebault_trophic_2005]. Vertical diversity, as network structure, can have
strong effects on temporal stability. However, there are few assesments of the
effects of horizontal and diversity on temporal stability in natural
communities.

#### Stream fish communities is a good model to study effects of horizontal & vertical diversity

Stream fish communities constitutes good model to study the joint effects of
horizontal & vertical diversity on total biomass and the temporal stability of
biomass. Despite that fish trophic interactions are not recorded in nature, a
consequent body of litterature supports that they can be infered through body
size and diet informations [@gravel_inferring_2013; @jennings_weak_2001;
@brose_predator_2019]. Most of the biomass-diversity and
stability-diversity relationships have been assessed in terrestrial ecosytems
while there is a evident need to test such relationships in a variety of
ecosystem types. Stream ecosystems are caracterized by the flowing water and
depends for the large part from the input of terrestrial ecosystems
. Furthermore, the biotic activity in stream ecosystems
determines for a large part the water quality [@cummins_structure_1974] and thus  
are crucial to maintain **human selfish goals (I can do better than that)**.

#### Environment plays a big role in structuring communities

Environment also largely structure communities and is important to take in
account when considering natural communities. Environment directly affects
organism performance, and subsequently, abundance of species and in turn
determines total biomass [Add general ref; @grace_integrative_2016] and temporal
stability [@de-boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @fried-peterson_drivers_2020]. Hydrology of streams
constraints the composition of fish communities [REF toulouse à balle], and is
also related to fish population temporal stability [@franssen_annual_2011].
However it is not clear which is the relative strength of direct and indirect
effects of environment on total biomass and its temporal stability (Figure
\@ref(fig:matmet) A).

#### SEM is perfect and our metamodel too

To reveal the direct and indirect effects of environment, horizontal & vertical
diversity on both total biomass and its temporal stability, we used Structural
Equation Modeling (SEM). It is one promising way to go beyond bivarite
relationships and to investigate complex relationships, direct and indirect,
between variables  [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. SEM makes possible to set
causal relationships between variables. This step can be critical, as the
estimated parameters can be determined by both the choice of the variables and
the links between them.  Therefore, the set of variables and links shoud be
supported by theory and litterature. We built a metamodel (set of variables and
links between them). Basically, we tested the links between environmental
conditions, horizontal & vertical diversity and both total biomass and its temporal
stability (Figure \@ref(fig:matmet) A). We considered that environment can
directly affect horizontal and vertical diversity, as well as total biomass and
its temporal stability. Horizontal diversity, measured as species richness
directly affects vertical diversity, total biomass and
its temporal stability. Food web structure was shown to be constrained by
species richness. In turn vertical diversity affects total biomass and
its temporal stability (Figure \@ref(fig:matmet) A). Two other questions can
only be answered when considering both direct & indirect relationships. (1) What
is the magnitude of environmental effects on total biomass and
its temporal stability compared to horizontal and vertical diversity? (2) Which
is the pathway of environment, horizontal & vertical diversity on total biomass
and its temporal stability ? (2a) For environment, is it mainly (i) direct,
indirect (ii) through horizontal diversity, (iii) through horizontal & vertical
diversity or (iv) indirect through only vertical diversity? (2b) For horizontal
diversity, is it mainly (i) direct, or (ii) indirect through vertical diversity?
Finally for temporal stability, how do environement, horizontal & vertical
diversity affect synchrony & species stability ?

We used an standardized monitoring of fish communities in french streams. 99
sites were followed more than 10 years on the 1995-2018 period. We inferred
(**XXX**) potential interaction networks corresponding to each sampling event,
based on trait matching. Horizontal diversity was measured as species richness
while vertical diversity was as connectance and average trophic level.
Environment was characterized by the stream size, temperature, flow,
enrichment, altitude and their stability (expect for altitude). Finally total
biomass was measured as median total biomass over the sampling period while its
temporal stability was quantified as the inverse of the CV.

Mid point used in niche model to characterize feeding links [@dunne__2006]

- Q1: diversity-stability, diversity-biomass in aquatic systems
- Q2: Effect of horizontal and vertical diversity on total biomass and stablity
- Q3: Relative contribution of environment and community structure to biomass
  and stability

@@
